Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-12-20 Manuscript/manuscriptPSsuppl.tex
Update references added in recent commit https://github.com/NMRLipids/NMRlipidsIVotherHGs/commit/a3978044eb65f6c1160abefce17e3a8ed7e1d7b1

The POPC bilayer consisting of a total of 128 lipids was constructed from single lipid structure taken 
from Ref.~\cite{10.1016/j.dib.2015.09.013}. The bilayer was hydrated by a total of 5120 water 
molecules, corresponding to 40 water molecules per lipid.

The MacRog lipid parameters \cite{10.1016/j.chemphyslip.2015.07.002}, obtained from 
\cite{10.1016/j.dib.2015.09.013}, were used, along with the TIP3P water model 
\cite{10.1063/1.445869}. The simulation parameters were identical to those used for 
POPC/POPC mixtures with the MacRog force field, see below. The simulation was run for 200~ns, 
out of which \hl{??~ns: (this is not tabulated in the MS or the SI)} were used for analyses.

@@

This membrane was further simulated after replacing Na$^+$ counterions by K$^+$ ones, which 
were described by the \AA{}qvist parameters \cite{10.1021/j100384a009}. Other topologies
were unchanged. The simulation parameters were the same as used for the POPC/POPS mixtures with 
the MacRog force field, see below. The simulation was run for 200~ns, out of which 150~ns was used 
in the analyses.

@@

The bilayers containing a total of 120 POPC and 24 POPS molecules distributed evenly among the 
two leaflets were set up using CHARMM-GUI \cite{10.1002/jcc.20945}.
The bilayers were solvated by 5760 water molecules, corresponding to 40 water molecules for lipid.
Additionally, 24 potassium ions were added to neutralize the charge of the POPS head groups.

@@

The lipids were described using the MacRog model \cite{10.1021/jp5016627,10.1016/j.chemphyslip.2015.07.002,10.1016/j.dib.2016.03.067}. The POPC topologies were obtained from 
Ref.~\citenum{10.1016/j.dib.2015.09.013}, whereas the POPS topology was adapted from 
Ref.~\citenum{10.1016/j.dib.2016.03.067}. Notably, the chains were reversed in order to modify 
OPPS into POPS. Unlike the D stereoisomer present in Ref.~\cite{10.1016/j.dib.2016.03.067}, the 
POPS head group conformations generated by CHARMM-GUI are in the naturally occurring L stereoisomer.
TIP3P model \cite{10.1063/1.445869} was used for water, and the \AA{}qvist parameters 
\cite{10.1021/j100384a009}, standard for the OPLS force field, were employed for the ions.

The simulation parameters were taken from Ref.~\citenum{10.1016/j.dib.2016.03.067}. 
Periodic boundary conditions were employed in all dimensions. The list of neighbors was updated every step 
\hl{I used nstlist=1 because it's used in the mdp in the article quoted. In DOI: 10.1021/jp5016627, nstlist=10 is used}
The Lennard-Jones interactions were cut off at 1~nm, whereas the smooth particle mesh Ewald algorithm 
\cite{10.1063/1.470117} was used for electrostatics. Dispersion correction \cite{10.1021/jp0735987} 
was applied to both energy and pressure. Temperature was kept constant at 298~K using the 
Nos\'{e}--Hoover thermostat \cite{10.1063/1.447334,10.1103/PhysRevA.31.1695}. The lipids and 
the solvent were coupled separately, and a time constant of 0.4~ps was employed. Pressure was 
maintained semi-isotropically at 1~bar using the Parrinello--Rahman barostat with a time constant of 1~ps 
\cite{10.1063/1.328693}. All non-water bonds were constrained with the LINCS algorithm 
\cite{10.1002/(SICI)1096-987X(199709)18:12<1463::AID-JCC4>3.0.CO;2-H,10.1021/ct700200b},
whereas the bonds in water molecules were constrained using SETTLE \cite{10.1002/jcc.540130805}.
The integration time step was set to 2~fs, and the simulations were performed either for 400~ns 
(only counterions), 300~ns (systems with KCl), or 600~ns (systems with CaCl$_2$). The last 
250, 200, or 300~ns of the simulations were included in the analyses, respectively.

>>>>

The POPC bilayer consisting of a total of 128 lipids was constructed from single lipid structure taken 
from Ref.~\cite{kulig15c}. The bilayer was hydrated by a total of 5120 water 
molecules, corresponding to 40 water molecules per lipid.

The MacRog lipid parameters \cite{kulig15b}, obtained from 
\cite{kulig15c}, were used, along with the TIP3P water model 
\cite{jorgensen83}. The simulation parameters were identical to those used for 
POPC/POPC mixtures with the MacRog force field, see below. The simulation was run for 200~ns, 
out of which \hl{??~ns: (this is not tabulated in the MS or the SI)} were used for analyses.

@@

This membrane was further simulated after replacing Na$^+$ counterions by K$^+$ ones, which 
were described by the \AA{}qvist parameters \cite{aqvist90}. Other topologies
were unchanged. The simulation parameters were the same as used for the POPC/POPS mixtures with 
the MacRog force field, see below. The simulation was run for 200~ns, out of which 150~ns was used 
in the analyses.

@@

The bilayers containing a total of 120 POPC and 24 POPS molecules distributed evenly among the 
two leaflets were set up using CHARMM-GUI \cite{lee16,jo18}.
The bilayers were solvated by 5760 water molecules, corresponding to 40 water molecules for lipid.
Additionally, 24 potassium ions were added to neutralize the charge of the POPS head groups.

@@

The lipids were described using the MacRog model \cite{maciejewski14,kulig15b,rog16}. The POPC topologies were obtained from 
Ref.~\citenum{kulig15c}, whereas the POPS topology was adapted from 
Ref.~\citenum{rog16}. Notably, the chains were reversed in order to modify 
OPPS into POPS. Unlike the D stereoisomer present in Ref.~\cite{rog16}, the 
POPS head group conformations generated by CHARMM-GUI are in the naturally occurring L stereoisomer.
TIP3P model \cite{jorgensen83} was used for water, and the \AA{}qvist parameters 
\cite{aqvist90}, standard for the OPLS force field, were employed for the ions.

The simulation parameters were taken from Ref.~\citenum{rog16}. 
Periodic boundary conditions were employed in all dimensions. The list of neighbors was updated every step 
\hl{I used nstlist=1 because it's used in the mdp in the article quoted. In DOI: 10.1021/jp5016627, nstlist=10 is used}
The Lennard-Jones interactions were cut off at 1~nm, whereas the smooth particle mesh Ewald algorithm 
\cite{essman95} was used for electrostatics. Dispersion correction \cite{shirts07} 
was applied to both energy and pressure. Temperature was kept constant at 298~K using the 
Nos\'{e}--Hoover thermostat \cite{nose84,hoover85}. The lipids and 
the solvent were coupled separately, and a time constant of 0.4~ps was employed. Pressure was 
maintained semi-isotropically at 1~bar using the Parrinello--Rahman barostat with a time constant of 1~ps 
\cite{parrinello81}. All non-water bonds were constrained with the LINCS algorithm 
\cite{hess97,hess07},
whereas the bonds in water molecules were constrained using SETTLE \cite{miyamoto92}.
The integration time step was set to 2~fs, and the simulations were performed either for 400~ns 
(only counterions), 300~ns (systems with KCl), or 600~ns (systems with CaCl$_2$). The last 
250, 200, or 300~ns of the simulations were included in the analyses, respectively.
